---
title: "Estimating the future shift patterns of shallow-water and deep-sea Crustacea"
author: 'Marianna V. P. Simões, Hanieh Saeedi, Marlon E. Cobos & Angelika Brandt '
output: 
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
---

This file contains the R code used to perfom the pre-modeling, modeling and post-modeling of each deep-sea and shallow-water species ecological niche, included in the manuscript "Estimating the future shift patterns of shallow-water and deep-sea Crustacea" .A schematic figure of the entire modeling process can be found at the Extended Data Figure 1. 

The three modeling steps include:

A. Pre modeling: 1. Distribution data capture and cleaning from open-source databases (GBIF and OBIS); 2. Distribution data thinning;3. Organization of species folders; 4. Split of occcurence into train and test;5. Creation of calibration area ("M") 

B. Modeling: 1. Creation of candidate models; 2. Evaluation of candidate models;  3. Selecting best model base don aic and ROC; 4. Creation of Final models; 5. Transforming final models to binary.

C. Post modeling: 1. Projection changes; 2. Variance in model predictions; 3. Extrapolation analysis (MOP); 
4. Binary MOPs;  5. Subtracting MOP from Final models (post-MOP models); 6. Calculating centroid shift; 7. MOPs summary 
 

Most of analysis used the kuenm R package, to which a brief tutorial can be found in the "https://github.com/marlonecobos/kuenm" . There, you can also find instructions for the directory structure, necessary data and additional software for its installation. 

## Pre-modeling: 

The pre- modeling phase includes steps on the preparation of occurence data and environemntal data for the creation of ecological niche models. It includes: 

1. Distribution data capture and cleaning from open-source databases (GBIF and OBIS); 
2. Distribution data thinning;
3. Oragnization of species folders;
4. Split of occcurence into train and test;
5. Creation of calibration area ("M") 

### Distribution data:

Distribution data were based on four deep-sea expeditions to the NW Pacific including the Sea of Japan Biodiversity Study (SoJaBio; 2010) (Malyutina & Brandt 2013), Kuril-Kamchatka Biodiversity Study (KuramBio I and II; 2012 - 2016) (Brandt & Malyutina 2015), and the Sea of Okhotsk Biodiversity Study (SokhoBio; 2015)  (Malyutina et al. 2018). The beneficial project assembled these datasets The list of species created based on the cited databses, was used to extract data on the Ocean Biogeographic Information System (OBIS) (www.iobis.org) and Global Biodiversity Information System (GBIF).To classify the species according to the depth, we used the maximum depth associated with records in the used databases. 

Occurence records were manually checked for suitability,  dubious records were corrected following protocol in Cobos et al. (2018).All species names were matched against the World Register of Marine Species (WoRMS) and synonyms reconciled (www.marinespecies.org). Finally, species were categorized into two groups: shallow-water (0-500m) and deep-sea (>500m), following classification by the World Register of Deep-Sea Species (WoRDSS). 

### Data capture: Obtaining occourence data from Ocean Biogeographic Information System (OBIS)

```{r, eval=FALSE}
sp <- read.csv("List of Crustacean Species.csv", stringsAsFactors = FALSE)
dim(sp)
name <- sp$ScientificName

spdata <- lapply(names, function(name) {
  
  message(name)
  
  return(robis::occurrence(name))
  
})

data <- dplyr::bind_rows(spdata)

```

### Data capture: Obtaining occourence data from Global Biodiversity Information System (GBIF). 

```{r, eval=FALSE}

spvector <- as.character(read.csv("List of Crustacean species.csv")[, 1])# binomial names

## Getting info species by species 
occ_count <- list() # object to save info on number of georeferenced records per species 

for (i in 1:length(spvector)) {
  sps <- try(name_lookup(query = spvector[i], rank = "species", 
                         return = "data", limit = 100), silent = TRUE) # information 
  #about the species
  
  sps_class <- class(sps)
  
  # avoiding errors from GBIF (e.g., species name not in GBIF)
  if(sps_class[1] == "try-error") {
    occ_count[[i]] <- c(Species = spvector[i], keys = 0, counts = 0) # species not in GBIF
    cat("species", spvector[i], "is not in the GBIF database\n")
    
  }else {
    keys <- sps$key # all keys returned
    counts <- vector() # object to save info on number of records per key
    
    for (j in 1:length(keys)) { # testing if keys return records
      counts[j] <- occ_count(taxonKey = keys[j], georeferenced = TRUE) 
    }
    
    if (sum(counts) == 0) { # if no info, tell the species
      occ_count[[i]] <- c(Species = spvector[i], keys = "all", counts = 0) # species not in GBIF
      cat("species", spvector[i], "has no goereferenced data\n")
      
    }else { # if it has info, use the key with more records, which is the most useful
      if (length(keys) == 1) { # if it is only one key
        key <- keys # detecting species key 
        occ_count[[i]] <- cbind(spvector[i], counts) # count how many records
        
      }else { # if its more than one key
        keysco <- cbind(keys, counts)
        keysco <- keysco[order(keysco[, 2]), ]
        key <- keysco[dim(keysco)[1], 1] # detecting species key that return information
        occ_count[[i]] <- c(Species = spvector[i], keysco[dim(keysco)[1], ])# count how many records
      }
      # getting the data from GBIF
      occ <- try(occ_search(taxonKey = key, return = "data", limit = 10000), silent = TRUE) 
      occ_class <- class(occ)
      
      # avoiding errors from GBIF
      while (occ_class[1] == "try-error") {
        occ <- try(occ_search(taxonKey = key, return = "data", limit = 10000), silent = TRUE) 
        occ_class <- class(occ)
        
        if(occ_class[1] != "try-error") {
          break()
        }
      }
      
      # following steps
      occ_g <- occ
      occ_g <- occ_g[, c(1, 2, 4, 3, 5:dim(occ_g)[2])] # reordering longitude and latitude
      
      # keeping only unique georeferenced records. 
      # excluding no georeferences
       # excluding duplicates
      occ_g <- occ_g[!is.na(occ_g$decimalLatitude) & !is.na(occ_g$decimalLongitude), ] 
      occ_g <- occ_g[!duplicated(paste(occ_g$name, occ_g$decimalLatitude,
                                       occ_g$decimalLongitude, sep = "_")), ]
      
      # writting file
      # csv file name per each species
      file_name <- paste(gsub(" ", "_", spvector[i]), "csv", sep = ".") 
      write.csv(occ_g, file_name, row.names = FALSE) # writing inside each genus folder
      
      cat(i, "of", length(spvector), "species\n") # counting species per genus 
    }
  }
}

genus_data <- do.call(rbind, occ_count) # making the list of countings a table
# making countings numeric
genus_data <- data.frame(genus_data[, 1], genus_data[, 2], as.numeric(genus_data[, 3])) 
names(genus_data) <- c("Species", "Key", "N_records") # naming columns

# writing the table
file_nam <- "Species_record_count.csv" # csv file name for all species
write.csv(genus_data, file_nam, row.names = FALSE) # writing inside each genus folder
 
```

### Distribution thinning: 

To avoid excessive clustering within any particular region, and to attain a uniform distribution of occurrences across the species’ known distribution, we explored different filtering distances in sequential steps with the package ‘spThin’ (Aiello-Lammens et al., 2015). Once a satisfactory distribution of occurrence data was obtained (i.e., when spatial clustering of records was no longer evident), we split the occurrences randomly in equal halves: one set of locations for calibration and another for evaluation during model calibration. Species with at least 30 records were retained and error-checked to meet appropriate standards for ecological niche modeling. 

```{r, eval=FALSE}

spvec <- dir()
exclude <- c("Variables", "G_variables")
spvector <- spvec[!spvec %in% exclude]

final_occ <- list()

for (i in 1:length(spvector)) {
  # reading species data
  sp <- read.csv(paste(spvector[i], "/", paste(spvector[i], ".csv", sep = ""), sep = ""))
  sp_nodup <- unique(sp)
  
  if (dim(sp_nodup)[1] > 1000) {
    sp_nodup <- sp_nodup[sample(1:nrow(sp_nodup), 1000), ]
  }
  
  # thinning
  thinnedcat_a <- thin( loc.data = sp_nodup, lat.col = "decimalLatitude", 
                        long.col = "decimalLongitude", 
                        spec.col = "species", thin.par = 10, reps = 5, 
                        locs.thinned.list.return = TRUE, write.files = FALSE, 
                        max.files = 1, out.dir = spvector[i], 
                        out.base = "cat_a_thined", write.log.file = FALSE )
  
  thin <- cbind(as.character(sp_nodup[1, 1]), thinnedcat_a[[5]])
  colnames(thin) <- colnames(sp_nodup)
  
  write.csv(thin, paste0(spvector[i], "/", paste0(spvector[i], "_thin.csv")),
            row.names = FALSE)
  
  # training and testing split
  thin$check <- paste(thin[,2], thin[,3], sep = "_") # column to make the split
  train <- thin[sample(nrow(thin), round((length(thin[,1]) / 4 * 3))), ] # training data
  test <- thin[!thin[,4] %in% train[,4],] # testing data
  
  thin$check <- NULL # deleting column to make splits
  train$check <- NULL
  test$check <- NULL
  # writing files (change names as needed)
  write.csv(thin, paste0(spvector[i], "/", paste0(spvector[i], "_joint.csv")),
            row.names = FALSE) 
  write.csv(train, paste0(spvector[i], "/", paste0(spvector[i], "_train.csv")),
            row.names = FALSE)
  write.csv(test, paste0(spvector[i], "/", paste0(spvector[i], "_test.csv")),
            row.names = FALSE)
  
  # number of final records per species
  final_occ[[i]] <- data.frame(species = spvector[i], n_records = dim(thin)[1])
  
  cat("\nSpecies", i, "of", length(spvector), "finished\n")
}

# writing summary of species number of records
final_n <- do.call(rbind, final_occ)
write.csv(final_n, "summary_occ_n.csv", row.names = FALSE)


```


### Organization of species folders

```{r, eval=FALSE}
x <- read.csv('CSV with all occurences for all species')
sp <- unique(x$Species) #str(sp)

n = 30 #number of records 
cols = 2:4 # columns extracted from dataset - species, long, lat
for (i in 1:length(sp)) {
  sptable <- x[x$Species == sp[i], cols]
  if (dim(sptable)[1] >= n) {
    dir.create(paste(sp[i], collapse = "_"))
    write.csv(sptable, paste(paste(sp[i], collapse = '_'), '\\',
                             paste(sp[i], collapse = '_'), ".csv", sep = ""), row.names = FALSE)
  }
  
}



```

### Split of occcurence into train and test

Once a satisfactory distribution of occurrence data was obtained (i.e., when spatial clustering of records was no longer evident), we split the occurrences randomly in equal halves: one set of locations for calibration and another for evaluation during model calibration

```{r, eval=FALSE}

spvec <- dir()
exclude <- c("Variables", "G_variables")

for (i in 1:length(spvector)) {
  all <- read.csv(paste(spvector[i], "/", paste(spvector[i], ".csv", sep = ""), sep = ""))
  all <- unique(all)
  
  
  all$check <- paste(all[,2], all[,3], sep = "_")
  train <- all[sample(nrow(all), round((length(all[,1])/4 *3))), ]
  test <- all[!all[,4] %in% train[,4], ]
  
  all$check <- NULL
  train$check <- NULL
  test$check <- NULL
  
  write.csv(all, paste(spvector[i], "/", paste(spvector[i], "_joint.csv", sep = ""), 
                       sep = ""), row.names = FALSE)
  write.csv(train, paste(spvector[i], "/", paste(spvector[i], "_train.csv", sep = ""),
                         sep = ""), row.names = FALSE)
  write.csv(test, paste(spvector[i], "/", paste(spvector[i], "_test.csv", sep = ""),
                        sep = ""), row.names = FALSE)
}

```
 
### Creation of calibration area ("M")

Calibration areas included a buffer of 2 decimal degrees from the occurrences of each species (‘M’; Barve et al., 2011)

```{r, eval=FALSE}

spvec <- dir()
exclude <- c("Variables", "G_variables")
spvector <- spvec[!spvec %in% exclude]

scenarios <- c("present")

var_list <- list(stack(list.files("Variables/set1/present", pattern = ".asc",
                                  full.names = TRUE)), 
                 stack(list.files("Variables/set2/present", pattern = ".asc", 
                                  full.names = TRUE)))

var_names <- list(list.files("Variables/set1/present", pattern = ".asc",
                             full.names = FALSE),
                  list.files("Variables/set2/present", pattern = ".asc",
                             full.names = FALSE))

sets <- c("set1", "set2")


for (i in 1:length(spvector)) {
  # reading species data
  sp <- read.csv(paste(spvector[i], "/", paste(spvector[i], ".csv", sep = ""), 
                       sep = ""))
  sp_nodup <- unique(sp)
  
  # create buffer
  WGS84 <- sp::CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
  occ_pr <- sp::SpatialPointsDataFrame(coords = sp_nodup[, 2:3], data = sp_nodup,
                                       proj4string = WGS84)
  
  buff_area <- rgeos::gBuffer(occ_pr, width = 2)
  #buff_areap <- raster::disaggregate(buff_area)
  
  # folder for Ms
  dir.create(paste(spvector[i], "M_variables", sep = "/"))
  
  for (j in 1:length(var_list)) {
    set <- var_list[[j]]
    
    # folder for sets
    infolder <- paste(spvector[i], "M_variables", sets[j], sep = "/")
    dir.create(infolder)
    
    # mask variables
    masked <- mask(crop(set, buff_area), buff_area)
    
    # write variables
    for (k in 1:length(unstack(masked))) {
      writeRaster(masked[[k]], filename = paste(infolder, var_names[[j]][k],
                                                sep = "/"), format = "ascii")
    }
    cat("   Set", j, "of", length(var_list), "finished\n")
  }
  
  cat("Species", i, "of", length(spvector), "finished\n")
}


```


## Modeling 

The modeling phase includes:

1. Creation of candidate models; 
2. Evaluation of candidate models;
3. Selecting best model base don aic and ROC;
4. Creation of Final models;
5. Transforming final models to binary.
 
 
 The creation of candidate models (kuenm_cal) concerns the creation of a set of  models that test a broad suites of parameter combinations, such as, distinct regularization multiplier values, various feature classes, and different sets of environmental variables. In this study we tested  126 candidate models per each species, with parameterizations resulted from the combinations of nine regularization multipliers (0.1–1.0 at intervals of 0.2, 2–5 at intervals of 1), seven feature classes representing combinations of linear, quadratic and product, two distinct sets of variables (without topography, ‘set 1’ and including topography, ‘set 2’), with projections created allowing extrapolation and clamping.  
 Following this step we evaluated the models (kuenm_ceval), to select candidate models and their associated parameters to identify the best models based on three distinct criteria: statistical significance (based on partial ROC analyses), prediction ability (we use omission rates, but other metrics, such as overall correct classification rate, can also be used), and model complexity (here evaluated using AICc).
 Selected model parameterizations were the ones that resulted in significant models (partial ROC with E = 5%, 500 iterations, and 50 percent of data for bootstrapping; Peterson et al. 2008), that had omission rates lower than a previously defined error rate (E = 5%, Anderson et al 2003), and the lowest AICc value per each species (Warren et al. 2010), in that order. 
 Then, after selecting parametrizations that produce best models, we created final models (kuenm_mod) for the present and globally projected to 2050 and 2100 for two representative concentration pathway emission scenarios (RCP 26 and 85). 

### Candidate models: creation and eveluation

```{r, eval=FALSE}

spvec <- dir()
exclude <- c("Variables", "G_variables")
spvector <- spvec[!spvec %in% exclude]
sp1 <- spvector

# occurrences
occ_joint <- paste(sp1, "/", paste(sp1, "_joint.csv", sep = ""), sep = "")
occ_tra <- paste(sp1, "/", paste(sp1, "_train.csv", sep = ""), sep = "")
occ_test <- paste(sp1, "/", paste(sp1, "_test.csv", sep = ""), sep = "")

# variables
M_var_dir <- paste(sp1, "/", "M_variables", sep = "")# Calibration area of each species
G_var_dir <- "G_variables" #projection area for all species

# maxent argument
batch_cal <- paste(sp1, "/", "Candidate_models1", sep = "")
out_dir <- paste(sp1, "/", "Candidate_Models1", sep = "")
reg_mult <- c(seq(0.5, 1, 0.5), seq(2, 2, 1))
f_clas <- "no.t.h"
background <- 10000
maxent_path <- "C:\\Maxent"
wait <- TRUE
run <- TRUE
out_eval <- paste(sp1, "/", "Calibration_results1", sep = "")
threshold <- 5
rand_percent <- 50
iterations <- 500
kept <- FALSE
selection <- "OR_AICc"
paral_proc <- FALSE
batch_fin <- paste(sp1, "/", "Final_models", sep = "")
mod_dir <- paste(sp1, "/", "Final_Models", sep = "")
rep_n <- 10
rep_type <- "Bootstrap"
jackknife <- FALSE
out_format <- "logistic"
project <- TRUE
ext_type <- "ext_clam"
write_mess <- TRUE
write_clamp <- FALSE
args <- NULL

# Creation of candidate models, evaluation of candidate models and final models
for (i in 1:length(occ_joint)) {
  # create candidate models
   kuenm_cal(occ.joint = occ_joint[i], occ.tra = occ_tra[i],
             M.var.dir = M_var_dir[i], batch = batch_cal[i],
             out.dir = out_dir[i], reg.mult = reg_mult, f.clas = f_clas,
             background = background,maxent.path = maxent_path,
             wait = wait, run = run)
  # evaluate all candidate models and select the best based 
  #on pROC, Omission rate and complexity AICc
  kuenm_ceval(path = out_dir[i], occ.joint = occ_joint[i], 
              occ.tra = occ_tra[i], occ.test = occ_test[i],
              batch = batch_cal[i], out.eval = out_eval[i], threshold = threshold,
              rand.percent = rand_percent, iterations = iterations, kept = kept,
            selection = selection, parallel.proc = paral_proc)
}
```

### Selecting best model parameters

```{r, eval=FALSE}

deep <- dir(pattern = "Done$", full.names = TRUE)
deep <- deep[2]

for (i in 1:length(deep)) {
  spvec <- dir(path = deep[i], full.names = TRUE)
  bar <- dir(path = deep[i], pattern= 'Variables', full.names = TRUE)
  g_var <- dir(path = deep[i], pattern= 'G_variables', full.names = TRUE)           
  extra <- dir(path = deep[i], pattern= 'Calibration', full.names = TRUE)  
  exclude <- c(bar, g_var, extra)
  spvector <- spvec[!spvec %in% exclude]
  
  for (j in 1:length(spvector)) {
    best <- list.files(path = paste(spvector[j], "Calibration_results1", sep = "/"),
                       pattern = "best", full.names = TRUE)
    bestr <- read.csv(best)
    
    file.rename(best, paste(spvector[j], "Calibration_results1", 
                            "tseb_candidate_models_OR_AICc.csv", sep = "/"))
    
    bestr <- bestr[bestr[, 6] == 0, ]
    
    if (dim(bestr)[2] > 1) {
      models <- bestr[, 1]
      
      sn <- ".*set2"
      stn <- gregexpr(sn, models)
      stan <- regmatches(models, stn)
      statn <- unlist(stan)
      
      if (length(statn) > 0) {
        bestb <- bestr[bestr[, 1] == statn, ][1, ]
      }else {
        bestb <- bestr[1, ]
      }
    }else {
      bestb <- bestr
    }
    
    write.csv(bestb, file = best, row.names = FALSE)
    cat("   species", j, "of", length(spvector),'\n')
  }
  
  cat("group", i, "of", length(deep),'\n')
}

```

### Creation of Final models

```{r, eval=FALSE}

for (i in 1:length(occ_joint)) {
   # create final models using the parameterizations selected before
   kuenm_mod(occ.joint = occ_joint[i], M.var.dir = M_var_dir[i], 
             out.eval = out_eval[i], batch = batch_fin[i],
             rep.n = rep_n, rep.type = rep_type, jackknife = jackknife,
             out.dir = mod_dir[i],out.format = out_format, 
             project = project, G.var.dir = G_var_dir, ext.type = ext_type,
             write.mess = write_mess, write.clamp = write_clamp, 
             maxent.path = maxent_path, args = args, wait = wait, run = run)
  cat("\n\nAnalyses for species", i, "of", length(occ_joint), "finished\n")
}

```

### Final Model threshold

```{r, eval=FALSE}

spvec <- dir()
exclude <- c("Variables", "G_variables")
spvector <- spvec[!spvec %in% exclude]
sp1 <- spvector[1:20] 
sp2 <- gsub(" ", "_", sp1)

sp_name <- sp2
fmod_dir <- paste(sp1, "Final_Models", sep = '/')
format <- "asc"
project <- TRUE
stats <- c("med", "range")
rep <- TRUE
scenarios <- c("present", "2050_rcp26", "2050_rcp85", "2100_rcp26", "2100_rcp85")
ext_type <- c("EC") 
out_dir <- paste(sp1, "Final_Model_Stats", sep = '/')
occ <- paste(sp1, paste(sp1, "joint.csv", sep = '_'), sep = "/")
thres <- 5
curr <- "present"
emi_scenarios <- c("rcp26", "rcp85")
t_periods <- c("2050", "2100")
out_dir1 <-  paste(sp1, "Projection_Changes", sep = '/')
split <- 200 
out_dir2 <- paste(sp1, "Variation_from_sources", sep = '/')
i=1
for (i in 1:length(sp1)) {
  if(i != 1)    {
     kuenm_modstats(sp.name = sp_name[i], fmod.dir = fmod_dir[i], 
                    format = format, project = project, statistics = stats,
                    replicated = rep, proj.scenarios = scenarios, 
                   ext.type = ext_type, out.dir = out_dir[i]) 
      }
}
```


## Post-modeling 
The post-modeling phase includes: 

 1. Projection changes; 
 2. Variance in model predictions;
 3. Extrapolation analysis (MOP); 
 4. Binary MOPs; 
 5. Subtracting MOP from Final models (post-MOP models);
 6. Calculating centroid shift
 
 After iobtaining the final models, we performed map algebra to represent how and where models projected in time change compared to the current one (kuenm_projchanges).Moreover, we model variation was assessed and represented geographically following the amount of variance in model predictions that came from replicates and emission scenarios (RCPs). 
 
 To assess the risks of strict extrapolation, we used the mobility-oriented parity metric (MOP; Owens et al. 2013), which consists of measuring similarity between the closest 5% of the environmental conditions of M to each environmental condition in the area of transference.  Areas with higher extrapolative values indicate higher uncertainty; and caution is required when interpreting likelihood of species presence in such areas (Alkishe et al. 2017). 
 Then, we threshold MOP results, considering only areas with zero similarity as strict extrapolation, and masked areas of extrapolation from binary final models, using binary MOPs, creating final binary models with no extrapolation (post-MOP projections). 
 
 
### Projection changes

```{r, eval=FALSE}
 for (i in 1:length(sp1)) {
    kuenm_projchanges(occ = occ[i], fmod.stats = out_dir[i], threshold = thres, 
                      current = curr, emi.scenarios = emi_scenarios,  
                      time.periods = t_periods, ext.type = ext_type, 
                      out.dir = out_dir1[i])
    
      cat("\n\nAnalyses for species", i, "of", length(sp1), "finished\n")
 }

```
### Variance in model predictions

```{r, eval=FALSE}
 for (i in 1:length(sp1)) {
        kuenm_modvar(sp.name = sp_name[i], fmod.dir = fmod_dir[i], 
                     replicated = rep, format = format, project = project,
                     current = curr, emi.scenarios = emi_scenarios, 
                     time.periods = t_periods, ext.type = ext_type, 
                     split.length = split, out.dir = out_dir2[i])
    
    cat("\n\nAnalyses for species", i, "of", length(sp1), "finished\n")
 }

```
 
### MOP analysis

```{r, eval=FALSE}
percent <- 3 
paral <- FALSE 
comp_each <- 3000 

for (i in 1) {
  spvec <- dir(path = deep[i], full.names = TRUE)
  exclude <- paste(deep[i], c("Variables", "G_variables"), sep = "/")
  spvector <- spvec[!spvec %in% exclude]
  
  G_var_dir <- paste(deep[i], "G_variables", sep = "/")
  
  for (j in 101:length(spvector)) {
    sets <- list.files(path = paste(spvector[j], "Calibration_results1", sep = "/"),
                       pattern = "best", full.names = TRUE)
    sets_var <- as.character(read.csv(sets)[1, 1])
    
    sets_var <- gsub("^M.*_", "", sets_var)
    M_var_dir <- paste(spvector[j], "M_variables", sep = "/")
    out_mop <- paste(spvector[j], "MOP_results", sep = "/")
    
    kuenm_mmop(G.var.dir = G_var_dir, M.var.dir = M_var_dir, 
               sets.var = sets_var, out.mop = out_mop,
               percent = percent, parallel = paral, 
               comp.each = comp_each)
    
    cat("   species", j, "of", length(spvector), "==========\n")
  }
  
  cat("group", i, "of", length(deep), "======================\n")
}

```

### Transforming MOPs into binary
```{r, eval=FALSE}
#MOP Processing and Percentage of area -------------------

setwd('D://MOPS_Ready//arctic_shallow')
dir()
##MOPs
mops <- "MOP_\\d\\S*.tif$"
set <- "set\\d"

files <- list.files(pattern = mops, full.names = T, recursive = T) #list MOPS
files

new_files <- gsub("MOP_results", "MOP_results_bin", files)
new_files <- gsub(".tif$", "_bin.tif", new_files)
new_files

SET <- gsub(mops, "", new_files)
MAIN <- gsub(paste0(set, "/", mops), "", new_files)  #Mop binary names

##Binary
setwd('D:\\MODELS_Ready\\arctic_shallow')
sps <- dir()
exclude <- c("Variables", "G_variables")  #exclude g_variables and variables
sps <- sps[!sps %in% exclude]

bins <- "binary.*if$"

bin_files <- list.files(pattern = bins, full.names = T, recursive = T)
bin_files

bins_exc <- "binary_c.*on.tif$"
bin_files_exc <- list.files(pattern = bins_exc, full.names = T, recursive = T)
bin_files <- bin_files[!bin_files %in% bin_files_exc]

ex <- seq(2, 360, 2) #Number of bin_files 2 bcs of binary_ and binary_current
ne <- seq(8, 360, 8) #Number of bin_files 8 bcs there are 8 fies from the same species and we only need one

exs <- ex[!ex %in% ne]
bin_files <- bin_files[-exs]

new_bins <- gsub("Pro.*", 'Reduced_binary/', bin_files)
new_bins <- gsub(".tif$", "_bin.tif", new_bins)
new_bins

pat <- '.*%_'
name <- gsub(pat, '', new_files)
new_bins <- paste0(new_bins, name)

MAIN1 <- paste0(sps, "/", "Reduced_binary")
SCEN <- paste0(MAIN1, rep(c("/rcp_26", "/rcp_85"), length(MAIN1)))

results <- list()

for (i in 106:length(new_files)) {
  setwd("D:\\MOPS_Ready\\arctic_shallow")
  mop <- raster(files[i])
  setwd('D:\\MODELS_Ready\\arctic_shallow')
  bin <- raster(bin_files[i])
  
  mop <- mop > 0 # strict extrapolation
  #dir.create(MAIN[i]) #creating folder to save binary version of MOPS - MOP_results_bin
  #dir.create(SET[i])
  writeRaster(mop, filename = new_files[i], format = "GTiff", overwrite=TRUE)
  
  
  bin1 <- bin * mop
  #creating folder to save reduced version of MOPs - MOP - BINARY PREDIC
  dir.create(MAIN1[i])
  #dir.create(SCEN[i])
  writeRaster(bin1, filename = new_bins[i], format = "GTiff", overwrite=TRUE)
  
  
  # before
  vals <- na.omit(values(bin))
  uval <- sort(unique(vals))
  
  ## processing
  all <- length(vals)
  percents <- sapply(1:length(uval), function(x) {
    per <- round(sum(vals == uval[x]) * 100 / all, 2)
    return(per)
  })
  names(percents) <- uval
  
  # after
  vals <- na.omit(values(bin1))
  uval <- sort(unique(vals))
  
  ## processing
  all <- length(vals)
  percents1 <- sapply(1:length(uval), function(x) {
    per <- round(sum(vals == uval[x]) * 100 / all, 2)
    return(per)
  })
  names(percents1) <- uval
  
  results[[i]] <- list(before = percents, after = percents1)
  
  cat(i, "of", length(new_files), "\n")
}

names(results) <- new_bins

77m[[1]]$`before`

#cbind all befores 
#cbind afters 

library(plyr) 
df <- do.call("rbind", lapply(results, as.data.frame)) 
write.csv(df, file = "results_104.csv")

```


###  Calculating centroid shift

```{r, eval=FALSE}

#Calling New Binary (Binary - MOP)

devtools::install_github("marlonecobos/ellipsenm")
library(ellipsenm)
library(raster)

setwd('D:\\MODELS_Ready\\nw_arctic_deep')
sps <- dir()
exclude <- c("Variables", "G_variables")  #exclude g_variables and variables
sps <- sps[!sps %in% exclude]

var_list <- list(stack(list.files("D:/nw_shallow/Variables/set1/present",
                                  pattern = ".asc", full.names = TRUE)))

vars <- raster::stack(list.files("D:/nw_shallow/Variables/set1/present",
                                 pattern = ".asc", full.names = TRUE))

#Binary models - MOPS - future ones 
fut_new_bin <- list.files(pattern = '^2.*bin.tif$', full.names = T, recursive = T) 
fut_new_bin
#Binary models - MOPS - future ones 
fut_new_bin_ex <- list.files(pattern = '^2.*bin.tif2.*bin.tif$', full.names = T, recursive = T) 
fut_new_bin <- fut_new_bin[!fut_new_bin %in% fut_new_bin_ex]

#ex <- seq(1, 455, 2) #change the middle number for the number of "bin_files"
#ne <- seq(2, 455, 2)

#exs <- ex[!ex %in% ne]
#fut_new_bin <- fut_new_bin[-exs]
#fut_new_bin <- fut_new_bin[-231]
#fut_new_bin <- fut_new_bin[-232]

#Binary models - MOPS - Present ones
pres_new_bin <- list.files(pattern = '^present_bin.tif$', full.names = T, recursive = T)  
pres_new_bin

sp_occ <- list.files(pattern = 'joint.csv$', full.names = T, recursive = T)
sp_occ

pres_centroid <- list()
fut_centroid <- list()

for (i in 1:length(pres_new_bin)) {
  
  sp1 <- read.csv(sp_occ[i])
  sub1 <- subset(sp1, sp1$decimalLongitude < -180, sp1$decimalLongitude > 180) 
  sub2 <- subset(sp1, sp1$decimalLatitude < -90, sp1$decimalLatitude > 90) 
  sp2 <- sp1[ !(sp1 %in% sub1), ]
  sp <- sp2[ !(sp2 %in% sub2), ]
  
  pol <- concave_area(data = sp, longitude = "decimalLongitude", latitude = "decimalLatitude", 
                      length_threshold = 25, buffer_distance = 200)
  
  pres <- raster::raster(pres_new_bin[i])
  pres <- raster::mask(crop(pres,pol), pol)
  pres[pres[] == 0] <- NA
  pres_centroid[[i]] <- colMeans(xyFromCell(pres, which(pres[] == 1)))
  cents <- list()
  dist <- vector()
  NS <- vector()
  
  ifut <- ifelse(i == 1, 1, ((i - 1) * 4 + 1))
  lfut <- i * 4
  
  f_new_bin <- fut_new_bin[ifut:lfut]
  
  for (j in 1:length(f_new_bin)) {
    
    fut <- raster::raster(f_new_bin[j]) #calling present raster and transforming to poly
    fut <- raster::mask(crop(fut, pol), pol)
    fut[fut[] == 0] <- NA
    cents[[j]] <- colMeans(xyFromCell(fut, which(fut[] == 1)))
    
    #Distance from centroids
    
    dist[j] <-  pointDistance(pres_centroid[[i]], cents[[j]], lonlat=T)/1000 #distance in Km
    NS[j] <- ifelse(pres_centroid[[i]][2] > cents[[j]][2], 'South', 'North')
    
  }
  
  fut_centroid[[i]] <- data.frame(paste0(pres_new_bin[i], f_new_bin), 
                                  matrix(pres_centroid[[i]], nrow = 1), 
                                  do.call(rbind, cents), dist, NS)
  cat(i, 'of', length(pres_new_bin), '\n')
}

result <- do.call(rbind, fut_centroid)
```

###  MOP analysis summary

```{r, eval=FALSE}
library(raster)
setwd("D:/Black_HD/MOPS_Ready")

#####
# Changes in continuous suitabilities
### Deep
dir(pattern = "deep")

dirs <- c("arctic_deep", "nw_arctic_deep", "nw_deep") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "MOP.*tif$", 
                          full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)
ncl <- gregexpr(".*present.*", cont); ncla <- regmatches(cont, ncl); present <- unlist(ncla)

c5026 <- raster(y50_r26[1]) == 0
c5085 <- raster(y50_r85[1]) == 0
c10026 <- raster(y100_r26[1]) == 0
c10085 <- raster(y100_r85[1]) == 0
cpresent <- raster(present[1]) == 0

for (j in 1:(length(y50_r26) - 1)) {
  c5026 <- c5026 + (raster(y50_r26[j + 1]) == 0)
  c5085 <- c5085 + (raster(y50_r85[j + 1]) == 0)
  c10026 <- c10026 + (raster(y100_r26[j + 1]) == 0)
  c10085 <- c10085 + (raster(y100_r85[j + 1]) == 0)
  cpresent <- cpresent + (raster(present[j + 1]) == 0)
  cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
}

#c5026 <- c5026 / length(y50_r26)
#c5085 <- c5085 / length(y50_r26)
#c10026 <- c10026 / length(y50_r26)
#c10085 <- c10085 / length(y50_r26)
#cpresent <- cpresent / length(y50_r26)

dir.create("MOP_sum")

writeRaster(c5026, filename = "MOP_sum/sum_deep_MOP_5026.tif", format = "GTiff")
writeRaster(c5085, filename = "MOP_sum/sum_deep_MOP_5085.tif", format = "GTiff")
writeRaster(c10026, filename = "MOP_sum/sum_deep_MOP_10026.tif", format = "GTiff")
writeRaster(c10085, filename = "MOP_sum/sum_deep_MOP_10085.tif", format = "GTiff")
writeRaster(cpresent, filename = "MOP_sum/sum_deep_MOP_present.tif", format = "GTiff")


### Shallow
dir(pattern = "shallow")

dirs <- c("arctic_shallow", "nw_arctic_shallow", "nw_shallow") # are these all?


cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "MOP.*tif$", full.names = TRUE, 
                          recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)
ncl <- gregexpr(".*present.*", cont); ncla <- regmatches(cont, ncl); present <- unlist(ncla)

c5026 <- raster(y50_r26[1]) == 0
c5085 <- raster(y50_r85[1]) == 0
c10026 <- raster(y100_r26[1]) == 0
c10085 <- raster(y100_r85[1]) == 0
cpresent <- raster(present[1]) == 0

for (j in 1:(length(y50_r26) - 1)) {
  c5026 <- c5026 + (raster(y50_r26[j + 1]) == 0)
  c5085 <- c5085 + (raster(y50_r85[j + 1]) == 0)
  c10026 <- c10026 + (raster(y100_r26[j + 1]) == 0)
  c10085 <- c10085 + (raster(y100_r85[j + 1]) == 0)
  cpresent <- cpresent + (raster(present[j + 1]) == 0)
  cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
}

#c5026 <- c5026 / length(y50_r26)
#c5085 <- c5085 / length(y50_r26)
#c10026 <- c10026 / length(y50_r26)
#c10085 <- c10085 / length(y50_r26)
#cpresent <- cpresent / length(y50_r26)

writeRaster(c5026, filename = "MOP_sum/sum_shallow_MOP_5026.tif", format = "GTiff")
writeRaster(c5085, filename = "MOP_sum/sum_shallow_MOP_5085.tif", format = "GTiff")
writeRaster(c10026, filename = "MOP_sum/sum_shallow_MOP_10026.tif", format = "GTiff")
writeRaster(c10085, filename = "MOP_sum/sum_shallow_MOP_10085.tif", format = "GTiff")
writeRaster(cpresent, filename = "MOP_sum/sum_shallow_MOP_present.tif", format = "GTiff")


###  Processing data:

#Processing Model Change : 
library(raster)
setwd("D:/MODELS_Ready")

#####
# Changes in continuous suitabilities
### Deep
dir(pattern = "deep")

dirs <- c("arctic_deep", "nw_arctic_deep", "nw_deep") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "continuous_comparison.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

c5026 <- raster(y50_r26[1])
c5085 <- raster(y50_r85[1])
c10026 <- raster(y100_r26[1])
c10085 <- raster(y100_r85[1])

for (j in 1:(length(y50_r26) - 1)) {
  c5026 <- c5026 + raster(y50_r26[j + 1])
  c5085 <- c5085 + raster(y50_r85[j + 1])
  c10026 <- c10026 + raster(y100_r26[j + 1])
  c10085 <- c10085 + raster(y100_r85[j + 1])
  cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
}

c5026 <- c5026 / length(y50_r26)
c5085 <- c5085 / length(y50_r26)
c10026 <- c10026 / length(y50_r26)
c10085 <- c10085 / length(y50_r26)

dir.create("Model_changes")

writeRaster(c5026, filename = "Model_changes/cont_suitchan_deep_5026_avg.tif", format = "GTiff")
writeRaster(c5085, filename = "Model_changes/cont_suitchan_deep_5085_avg.tif", format = "GTiff")
writeRaster(c10026, filename = "Model_changes/cont_suitchan_deep_10026_avg.tif", format = "GTiff")
writeRaster(c10085, filename = "Model_changes/cont_suitchan_deep_10085_avg.tif", format = "GTiff")


### Shallow
dir(pattern = "shallow")

dirs <- c("arctic_shallow", "nw_arctic_shallow", "nw_shallow") # are these all?


cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "continuous_comparison.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

c5026 <- raster(y50_r26[1])
c5085 <- raster(y50_r85[1])
c10026 <- raster(y100_r26[1])
c10085 <- raster(y100_r85[1])

for (j in 1:(length(y50_r26) - 1)) {
  c5026 <- c5026 + raster(y50_r26[j + 1])
  c5085 <- c5085 + raster(y50_r85[j + 1])
  c10026 <- c10026 + raster(y100_r26[j + 1])
  c10085 <- c10085 + raster(y100_r85[j + 1])
  cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
}

c5026 <- c5026 / length(y50_r26)
c5085 <- c5085 / length(y50_r26)
c10026 <- c10026 / length(y50_r26)
c10085 <- c10085 / length(y50_r26)

writeRaster(c5026, filename = "Model_changes/cont_suitchan_shallow_5026_avg.tif", format = "GTiff")
writeRaster(c5085, filename = "Model_changes/cont_suitchan_shallow_5085_avg.tif", format = "GTiff")
writeRaster(c10026, filename = "Model_changes/cont_suitchan_shallow_10026_avg.tif", format = "GTiff")
writeRaster(c10085, filename = "Model_changes/cont_suitchan_shallow_10085_avg.tif", format = "GTiff")

#####
# Changes in binary suitable areas, losses and gains
### Deep
dir(pattern = "deep")

dirs <- c("arctic_deep", "nw_arctic_deep", "nw_deep") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "binary_comparison.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

change <- c(1, 2) # gain = 1, loss = 2
nam <- c("_gain.tif", "_loss.tif")

for (i in 1:length(change)) {
  c5026 <- raster(y50_r26[1]) == change[i]
  c5085 <- raster(y50_r85[1]) == change[i]
  c10026 <- raster(y100_r26[1]) == change[i]
  c10085 <- raster(y100_r85[1]) == change[i]
  
  for (j in 1:(length(y50_r26) - 1)) {
    c5026 <- c5026 + (raster(y50_r26[j + 1]) == change[i])
    c5085 <- c5085 + (raster(y50_r85[j + 1]) == change[i])
    c10026 <- c10026 + (raster(y100_r26[j + 1]) == change[i])
    c10085 <- c10085 + (raster(y100_r85[j + 1]) == change[i])
    cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
  }
  
  writeRaster(c5026, filename = paste0("Model_changes/bin_suitchan_deep_5026", nam[i]), format = "GTiff")
  writeRaster(c5085, filename = paste0("Model_changes/bin_suitchan_deep_5085", nam[i]), format = "GTiff")
  writeRaster(c10026, filename = paste0("Model_changes/bin_suitchan_deep_10026", nam[i]), format = "GTiff")
  writeRaster(c10085, filename = paste0("Model_changes/bin_suitchan_deep_10085", nam[i]), format = "GTiff")
  
  cat("changes", i, "of", length(change), "\n")
}


### Shallow
dir(pattern = "shallow")

dirs <- c("arctic_shallow", "nw_arctic_shallow", "nw_shallow") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "binary_comparison.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)

ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

change <- c(1, 2) # gain = 1, loss = 2
nam <- c("_gain.tif", "_loss.tif")

for (i in 1:length(change)) {
  c5026 <- raster(y50_r26[1]) == change[i]
  c5085 <- raster(y50_r85[1]) == change[i]
  c10026 <- raster(y100_r26[1]) == change[i]
  c10085 <- raster(y100_r85[1]) == change[i]
  
  for (j in 1:(length(y50_r26) - 1)) {
    c5026 <- c5026 + (raster(y50_r26[j + 1]) == change[i])
    c5085 <- c5085 + (raster(y50_r85[j + 1]) == change[i])
    c10026 <- c10026 + (raster(y100_r26[j + 1]) == change[i])
    c10085 <- c10085 + (raster(y100_r85[j + 1]) == change[i])
    cat("\tsum", j, "of", (length(y50_r26) - 1), "\n")
  }
  
  writeRaster(c5026, filename = paste0("Model_changes/bin_suitchan_shallow_5026", nam[i]), format = "GTiff")
  writeRaster(c5085, filename = paste0("Model_changes/bin_suitchan_shallow_5085", nam[i]), format = "GTiff")
  writeRaster(c10026, filename = paste0("Model_changes/bin_suitchan_shallow_10026", nam[i]), format = "GTiff")
  writeRaster(c10085, filename = paste0("Model_changes/bin_suitchan_shallow_10085", nam[i]), format = "GTiff")
  
  cat("changes", i, "of", length(change), "\n")
}


#Processing Variability found in ENMs
library(raster)
setwd("D:/MODELS_Ready")

#####
# Changes in continuous suitabilities
### Deep
dir(pattern = "deep")

dirs <- c("arctic_deep", "nw_arctic_deep", "nw_deep") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "var_emi_scenarios.tif$", full.names = TRUE, recursive = TRUE)
}
cont1 <- list()
for (i in 1:length(dirs)) {
  cont1[[i]] <- list.files(path = dirs[i], pattern = "0_var_replicates.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)
cont1 <- unlist(cont1)
cont <- list(cont, cont1)
nam <- c("_emi_scenarios.tif", "_replicates.tif")

for (i in 1:length(cont)) {
  ncl <- gregexpr(".*2050_var.*", cont[[i]]); ncla <- regmatches(cont[[i]], ncl); y50 <- unlist(ncla)
  ncl <- gregexpr(".*2100_var.*", cont[[i]]); ncla <- regmatches(cont[[i]], ncl); y100 <- unlist(ncla)
  
  c50 <- raster(y50[1])
  c100 <- raster(y100[1])
  
  for (j in 1:(length(y50) - 1)) {
    c50 <- c50 + raster(y50[j + 1])
    c100 <- c100 + raster(y100[j + 1])
    cat("\tsum", j, "of", (length(y50) - 1), "\n")
  }
  
  c50 <- c50 / length(y50)
  c100 <- c100 / length(y100)
  
  writeRaster(c50, filename = paste0("Model_changes/var_deep_2050_avg", nam[i]), format = "GTiff")
  writeRaster(c100, filename = paste0("Model_changes/var_deep_2100_avg", nam[i]), format = "GTiff") 
  cat("source", i, "of", length(cont), "\n")
}


### Shallow
dir(pattern = "shallow")

dirs <- c("arctic_shallow", "nw_arctic_shallow", "nw_shallow") # are these all?

cont <- list()
for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "var_emi_scenarios.tif$", full.names = TRUE, recursive = TRUE)
}
cont1 <- list()
for (i in 1:length(dirs)) {
  cont1[[i]] <- list.files(path = dirs[i], pattern = "0_var_replicates.tif$", full.names = TRUE, recursive = TRUE)
}

cont <- unlist(cont)
cont1 <- unlist(cont1)
cont <- list(cont, cont1)
nam <- c("_emi_scenarios.tif", "_replicates.tif")

for (i in 1:length(cont)) {
  ncl <- gregexpr(".*2050_var.*", cont[[i]]); ncla <- regmatches(cont[[i]], ncl); y50 <- unlist(ncla)
  ncl <- gregexpr(".*2100_var.*", cont[[i]]); ncla <- regmatches(cont[[i]], ncl); y100 <- unlist(ncla)
  
  c50 <- raster(y50[1])
  c100 <- raster(y100[1])
  
  for (j in 1:(length(y50) - 1)) {
    c50 <- c50 + raster(y50[j + 1])
    c100 <- c100 + raster(y100[j + 1])
    cat("\tsum", j, "of", (length(y50) - 1), "\n")
  }
  
  c50 <- c50 / length(y50)
  c100 <- c100 / length(y100)
  
  writeRaster(c50, filename = paste0("Model_changes/var_shallow_2050_avg", nam[i]), format = "GTiff")
  writeRaster(c100, filename = paste0("Model_changes/var_shallow_2100_avg", nam[i]), format = "GTiff") 
  cat("source", i, "of", length(cont), "\n")
}

#Variables contribution and permutation relevance to final models

#directory
setwd("D:/Black_HD/MODELS_Ready")

#subdirectories
dirs <- c("arctic_deep", "arctic_shallow", "nw_arctic_deep", "nw_arctic_shallow",
          "nw_deep", "nw_shallow")

exc <- paste0("./", dirs, "/G_variables", "/Variables")
i=1
# extracting by big groups
groups <- list()
for (i in 1:length(dirs)) {
  sps <- dir(path = dirs[i], full.names = TRUE)
  sps <- sps[!sps %in% exc]
  res <- list()
  for (j in 1:length(sps)) {
    mod <- dir(paste0(sps[j], "/Final_Models"), full.names = TRUE)[1]
    tor <- paste0(mod, "/maxentResults.csv")
    test <- read.csv(tor)
    nms <- colnames(test)
    test <- test[, c(grep(pattern = "contribution", x = nms),
                     grep(pattern = "permutation.importance", x = nms))]
    res[[j]] <- apply(test, 2, median)
  }
  groups[[i]] <- res
  
  cat("group", i, "of", length(dirs), "\n")
}

# dividing by major groups
str(deep)
deep <- c(groups[[1]], groups[[3]], groups[[5]])
shallow <- c(groups[[2]], groups[[4]], groups[[6]]) 

## four or three variables
deep3 <- deep[which(sapply(deep, function(x) {length(x) == 6}))]
deep4 <- deep[which(sapply(deep, function(x) {length(x) == 8}))]

shallow4 <- shallow[which(sapply(shallow, function(x) {length(x) == 8}))]
shallow5 <- shallow[which(sapply(shallow, function(x) {length(x) == 10}))]

# statistics
## three variables
deep3s <- apply(do.call(rbind, deep3), 2, median)
deep4s <- apply(do.call(rbind, deep4), 2, median)

shallow4s <-  apply(do.call(rbind, shallow4), 2, median)
shallow5s <-  apply(do.call(rbind, shallow5), 2, median)

# change variable names
cdeep3s <- deep3s[1:3]
names(cdeep3s) <- c("CVELOC", "Salinity", "Temperature")
pideep3s <- deep3s[4:6]
names(pideep3s) <- c("CVELOC", "Salinity", "Temperature")

cdeep4s <- deep4s[1:4]
names(cdeep4s) <- c("CVELOC", "Topo", "Salinity", "Temperature")
pideep4s <- deep4s[5:8]
names(pideep3s) <- c("CVELOC", "Topo", "Salinity", "Temperature")

cshallow4s <- shallow4s[1:4]
names(cshallow4s) <- c("CVELOC", "Ice", "Salinity", "Temperature")
pishallow4s <- shallow4s[5:8]
names(pishallow4s) <- c("CVELOC", "Ice", "Salinity", "Temperature")

cshallow5s <- shallow5s[1:5]
names(cshallow5s) <- c("CVELOC", "Topo", "Ice", "Salinity", "Temperature")
pishallow5s <- shallow5s[6:10]
names(pishallow5s) <- c("CVELOC", "Topo", "Ice", "Salinity", "Temperature")



## now for the others, remember that shallow and depp have distinct names and orders, check them before splitting and renaming

# plot (please fix this) contribution
barplot(cdeep3s, ylab = "Contribution", xlab = "Variables", ylim=c(0,60))
barplot(cdeep4s, ylab = "Contribution", xlab = "Variables", ylim=c(0,60))
barplot(cshallow4s, ylab = "Contribution", xlab = "Variables", ylim=c(0,60))
barplot(cshallow5s, ylab = "Contribution", xlab = "Variables", ylim=c(0,60))
## plot perutation importance
barplot(pideep3s, ylab = "Permutation importance", xlab = "Variables", ylim=c(0,60))
barplot(pideep4s, ylab = "Permutation importance", xlab = "Variables", ylim=c(0,60))
barplot(pishallow4s, ylab = "Permutation importance", xlab = "Variables", ylim=c(0,60) )
barplot(pishallow5s, ylab = "Permutation importance", xlab = "Variables", ylim=c(0,60))


#Correlation Between Deepth and species richeness
library(raster)
r <- raster('D:\\Black_HD\\MODELS_Ready\\Sum_binary_reduced_2100_85\\Shallow.tif')
e <- raster('D:\\Black_HD\\nw_shallow\\Variables\\set2\\2050_rcp85\\etoporst.asc')
d <- rasterToPoints( e )
v <- rasterToPoints( r )
head(v)
#write.csv(d, 'D:\\Review\\depth.csv')
#d <- read.csv('D:\\Review\\depth.csv')
write.csv(v, 'D:\\Review\\Shallow_2100_85.csv')
v <- read.csv('D:\\Review\\Shallow_2100_85.csv')

breaks <- seq(0,90, by=5)
spvector<-unique(breaks)
df1<-NULL
i=1

for (i in 1:length(spvector)) {
  sub <-v[v$y  >= spvector[i],]
  sub2 <-sub[sub$y <= spvector[i]+5,]
  mean <- mean(sub2$Shallow) # Number of records
  df1 = rbind(df1, data.frame(spvector[i],mean))
}


df_final6 <- cbind(df_final5, df1$mean)
dim(df_final6)
colnames(df_final6) <- c('lat', 'topo', 'Deep_2050_26', 'Shallow_2050_26', 'Deep_2050_85', 'shallow_2050_85', 'Deep_2100_26', 'Shallow_2100_26', 'Deep_2100_85', 'Shallow_2100_85')
write.csv(df_final6, 'D:\\Review\\CorDepth_SpRichenss.csv')
c <- read.csv('D:\\Review\\CorDepth_SpRichenss.csv')
a <- c[,2:10]
cor(a, method = 'pearson', use = 'pairwise.complete.obs')[1,-1]
install.packages("ggpubr")
library("ggpubr")
ggscatter(c, x = "topo", y = "Deep_2050_26", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Shallow_2050_26", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Deep_2050_85", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "shallow_2050_85", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Deep_2100_26", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Shallow_2100_26", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Deep_2100_85", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")
ggscatter(c, x = "topo", y = "Shallow_2100_85", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Depth", ylab = "Species Richeness")

```

#Preparing maps for each species - Suppm. Figures 20-21
#####
# Changes in binary suitable areas, losses and gains
### Deep
require(raster)
install.packages('rlang')
devtools::install_github("tidyverse/purrr")
library(purrr)
require(purrr)
require(raster)
require(maps)
require(mapproj)
setwd('E:\\Black_HD\\MODELS_Ready')
dir(pattern = "deep")
dirs <- c("arctic_deep", "nw_arctic_deep", "nw_deep") # are these all?

cont <- list()
spp <- list()

for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "binary_comparison.tif$", full.names = TRUE, recursive = TRUE)
  spp[[i]] <- list.dirs(path = dirs[i], recursive = FALSE, full.names = FALSE)
}
cont <- unlist(cont)

ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

spp <- gsub("_", " ", unlist(spp))

#### plotting
ras <- raster(y50_r26[1])
xlims <- extent(ras)[1:2]
ylims <- extent(ras)[3:4]

colmap <- c("#E5DFDF", "#DF3618", "#1724CA", "#9E9A9A")

pdf(file = "All_deep_project_changes.pdf", width = 14, height = 8)

layout(matrix(c(1:10, 11, 11), nrow = 4, byrow = TRUE), widths = c(1, 10, 10), 
       heights = c(1, 10, 10, 1))
j=1

for (j in 1:length(y50_r26)) {
  par(mar = rep(0, 4)); plot.new()
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "RCP 2.6", cex = 1.4)
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "RCP 8.5", cex = 1.4)
  
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "Year 2050", srt = 90, cex = 1.4)
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y50_r26[j]), col = colmap, axes = FALSE, add = TRUE)
  legend("topleft", legend = spp[j], inset = -0.01, bty = "n", cex = 1.2)
  box()
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y50_r85[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "Year 2100", srt = 90, cex = 1.4)
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y100_r26[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y100_r85[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  par(mar = rep(0, 4))
  plot.new()
  
  plot.new()
  legend("center", legend = c("Stable unsuitable", "Stable suitable", "Gain", "Loss"),
         fill = colmap[c(1, 4, 2, 3)], horiz = TRUE, bty = "n", cex = 1.3)
}

dev.off()


### Shallow
dir(pattern = "shallow")
dirs <- c("arctic_shallow", "nw_arctic_shallow", "nw_shallow") # are these all?

cont <- list()
spp <- list()

for (i in 1:length(dirs)) {
  cont[[i]] <- list.files(path = dirs[i], pattern = "binary_comparison.tif$", full.names = TRUE, recursive = TRUE)
  spp[[i]] <- list.dirs(path = dirs[i], recursive = FALSE, full.names = FALSE)
}

cont <- unlist(cont)
ncl <- gregexpr(".*2050/.*26.*", cont); ncla <- regmatches(cont, ncl); y50_r26 <- unlist(ncla)
ncl <- gregexpr(".*2050/.*85.*", cont); ncla <- regmatches(cont, ncl); y50_r85 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*26.*", cont); ncla <- regmatches(cont, ncl); y100_r26 <- unlist(ncla)
ncl <- gregexpr(".*2100/.*85.*", cont); ncla <- regmatches(cont, ncl); y100_r85 <- unlist(ncla)

spp <- gsub("_", " ", unlist(spp))

#### plotting
ras <- raster(y50_r26[1])
xlims <- extent(ras)[1:2]
ylims <- extent(ras)[3:4]

colmap <- c("#E5DFDF", "#DF3618", "#1724CA", "#9E9A9A")

pdf(file = "All_shallow_project_changes.pdf", width = 14, height = 8)

layout(matrix(c(1:10, 11, 11), nrow = 4, byrow = TRUE), widths = c(1, 10, 10), 
       heights = c(1, 10, 10, 1))

for (j in 1:length(y50_r26)) {
  par(mar = rep(0, 4)); plot.new()
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "RCP 2.6", cex = 1.4)
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "RCP 8.5", cex = 1.4)
  
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "Year 2050", srt = 90, cex = 1.4)
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y50_r26[j]), col = colmap, axes = FALSE, add = TRUE)
  legend("topleft", legend = spp[j], inset = -0.01, bty = "n", cex = 1.2)
  box()
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y50_r85[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  par(mar = rep(0, 4)); plot.new(); text(0.5, 0.5, "Year 2100", srt = 90, cex = 1.4)
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y100_r26[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  map(col = "transparent", xlim = xlims, ylim = ylims, mar = rep(0.5, 4))
  image(raster(y100_r85[j]), col = colmap, axes = FALSE, add = TRUE)
  box()
  
  par(mar = rep(0, 4))
  plot.new()
  
  plot.new()
  legend("center", legend = c("Stable unsuitable", "Stable suitable", "Gain", "Loss"),
         fill = colmap[c(1, 4, 2, 3)], horiz = TRUE, bty = "n", cex = 1.3)
}

dev.off()




## References 

```{r, eval=FALSE}
1. Brandt, A. & Malyutina, M. V. The German–Russian deep-sea expedition KuramBio 
(Kurile Kamchatka biodiversity studies) on board of the RV Sonne in 2012 following
the footsteps of the legendary expeditions with RV Vityaz. Deep Sea Research Part II: 
Topical Studies in Oceanography 111, 1–9 (2015).

2.	Malyutina, M. V. & Brandt, A. Introduction to sojabio (sea of japan biodiversity 
studies). Deep Sea Research Part II: Topical Studies in Oceanography 86–87, 1–9 (2013).

3.	Brandt, A. & Malyutina, M. V. The German–Russian deep-sea expedition KuramBio 
(Kurile Kamchatka biodiversity studies) on board of the RV Sonne in 2012 following
the footsteps of the legendary expeditions with RV Vityaz. Deep Sea Research Part II:
Topical Studies in Oceanography 111, 1–9 (2015).

4.	Malyutina, M. V., Chernyshev, A. V. & Brandt, A. Introduction to the SokhoBio 
(Sea of Okhotsk Biodiversity Studies) expedition 2015. Deep Sea Research Part II: 
Topical Studies in Oceanography 154, 1–9 (2018).

5.	Cobos, M. E., Jiménez, L., Nuñez-Penichet, C., Romero-Alvarez, D. & Simoes, M. 
Sample data and training modules for cleaning biodiversity information. Biodiv. Inf.
13, 49–50 (2018).

6.	Aiello-Lammens, M. E., Boria, R. A., Radosavljevic, A., Vilela, B. & Anderson, R. P. 
spThin: an R package for spatial thinning of species occurrence records for use in 
ecological niche models. Ecography 38, 541–545 (2015).

7.	Barve, N. et al. The crucial role of the accessible area in ecological niche modeling 
and species distribution modeling. Ecol. Modell. 222, 1810–1819 (2011).

8.	Peterson, A. T., Papeş, M. & Soberón, J. Rethinking receiver operating characteristic 
analysis applications in ecological niche modeling. Ecol. Modell. 213, 63–72 (2008).

9.	Anderson, R. P., Lew, D. & Peterson, A. T. Evaluating predictive models of species’ 
distributions: criteria for selecting optimal models. Ecol. Modell. 162, 211–232 (2003).

10.	Warren, D. L. & Seifert, S. N. Ecological niche modeling in Maxent: the importance of 
model complexity and the performance of model selection criteria. Ecol. Appl. 21, 335–342 
(2011).

11.	Owens, H. L. et al. Constraints on interpretation of ecological niche models by 
limited environmental ranges on calibration areas. Ecol. Modell. 263, 10–18 (2013).

12.	Alkishe, A. A., Peterson, A. T. & Samy, A. M. Climate change influences on the 
potential geographic distribution of the disease vector tick Ixodes ricinus. PLoS ONE 
12, e0189092 (2017).

```
